---
title: "p8105_hw1_unp2000"
output: github_document
---
# Problem 1: 
Create a data frame comprised of:
a random sample of size 8 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 8
a factor vector of length 8, with 3 different factor “levels”

```{r}
library(tidyverse)
hw1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = norm_samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h"),
  vec_factor = factor(c("small", "medium", "large", "medium", "small", "medium", "large", "large"))
)


##Try to take the mean of each variable in your dataframe. What works and what doesn’t?
mean(pull(hw1_df, norm_samp))
mean(pull(hw1_df, vec_logical))
mean(pull(hw1_df, vec_char))
mean(pull(hw1_df, vec_factor))
```
I am able to take the mean for my norm_sample (random sample of size 8 from a standard normal distribution) and vec_logical (logical vector indicating whether elements of the sample are greater than 0) variables because they do contain numeric data. I am unable to take the mean for vec_char (character vector of length 8) and vec_factor (factor vector length 8, with 3 different factor "levels") variables because they do not completely contain numeric data. 

```{r, eval = FALSE}
as.numeric(pull(hw1_df, vec_logical))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df,vec_factor))
```
The as.numeric function converts variables that were written in my logical, character, and factor vectors to numeric vectors. For my logical and factor vectors, this does explain what happens when we take the mean of both of these vectors. The logical vector converted the True and False answers to Numbers (0 and 1) and my factor vector was converted to numbers (3 = small, 2 = medium, 1 = large). The as.numeric function was unable to convert my character vector into a numeric vector because the variables listed were text. Because of this I was unable to take a mean. 

In a second code chunk:
convert the logical vector to numeric, and multiply the random sample by the result
convert the logical vector to a factor, and multiply the random sample by the result
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
```{r, eval = TRUE}
as.numeric(pull(hw1_df, vec_logical))*(pull(hw1_df,norm_samp))
as.factor(pull(hw1_df,vec_logical))*(pull(hw1_df,norm_samp))
as.numeric(as.factor(pull(hw1_df, vec_logical))*(pull(hw1_df,norm_samp)))
```

# Problem 2

This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

Create a data frame comprised of:
x: a random sample of size 500 from a standard Normal distribution
y: a random sample of size 500 from a standard Normal distribution
A logical vector indicating whether x + y > 1
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r}
prob_2_df = tibble(
  x = rnorm(500),
  y = rnorm (500),
  vec_logical = (x + y) >1, 
  vec_num_p2 = as.numeric(vec_logical),
  vec_factor_p2 = as.factor(vec_logical)
)
```


The size of the dataframe create above `prob_2_df` is `r nrow(prob_2_df)` rows and `r ncol(prob_2_df)` columns

The mean is `r mean(pull(prob_2_df,x))`.
The median is `r median(pull(prob_2_df,x))`.
The standard deviation is `r sd(pull(prob_2_df,x))`.
The proportion of cases for which x+y > 1 is `r mean(pull(prob_2_df,vec_logical))`.


```{r}
ggplot(prob_2_df) + geom_point(aes(x = x ,y = y), color = 'red')
ggsave("scatter_plot.pdf")

ggplot(prob_2_df) + geom_point(aes(x = vec_num_p2 ,y = y), color = 'green')

ggplot(prob_2_df) + geom_point(aes(x = vec_factor_p2 ,y = y), color = 'blue')
```
